R-LIME: Rectangular Constraints and Optimization for Local Interpretable Model-agnostic Explanation Methods

### 2024-02-16 paper.tex
Modify line feed code in paper.tex

This paper specifically focuses on \emph{local} \emph{model-agnostic} methods.
For instance, LIME (Local Interpretable Model-agnostic Explanations)
\cite{ribeiro2016why} approximates the complex model's decision boundary
locally using a simpler model.
Although LIME enables users to interpret the local behavior of complex models,
it does not optimize the approximation region \cite{radulovic2023bella} and
the effective scope of the explanation is not explicitly presented
\cite{ribeiro2018anchors}.

To address the limitations of LIME,
we propose a new method called R-LIME (Ruled LIME),
which interprets complex classifiers in an interpretable region.
R-LIME locally approximates a complex decision boundary linearly
in a rectangular region and maximizes the region as long as the accuracy of the linear
classifier is higher than a specified threshold.
The resulting rectangular region is interpretable for users because it is
expressed as a conjunction of feature predicates.

\section{Related Work}
\begin{figure}[t]
  \centering
  \vspace{0.5cm}
  \begin{subfigure}[t]{0.55\textwidth}
    \centering
    \includegraphics[width=\textwidth]{lime_vs_anchor_exp_a.png}
    \caption{Instances around the focal point.}\label{fig:lime_vs_anchor_exp_a}
    \vspace{0.5cm}
  \end{subfigure}
  \begin{subfigure}[t]{0.45\textwidth}
    \centering
    \includegraphics[width=\textwidth]{lime_vs_anchor_exp_b.png}
    \caption{%
      Explanations by LIME.
      They provide contributions of each feature to the output,
      but do not explicitly indicate its scope.
    }\label{fig:lime_vs_anchor_exp_b}
    \vspace{0.5cm}
  \end{subfigure}
  \begin{subfigure}[t]{0.55\textwidth}
    \centering
    \includegraphics[width=\textwidth]{lime_vs_anchor_exp_c.png}
    \caption{%
      Explanations by Anchor.
      They provide the effective scope of the explanation,
      but do not provide details about the influence of each feature.
    }\label{fig:lime_vs_anchor_exp_c}
  \end{subfigure}
  \caption{%
    Explanations by LIME and Anchor
    for an LSTM sentiment prediction model \cite{ribeiro2018anchors}.
  }\label{fig:lime_vs_anchor_exp}
\end{figure}
In this chapter,
we provide an overview of existing research on local model-agnostic explanations,
particularly focusing on studies closely related to our proposed method.

@@

\subsection{Quantitative Evaluation}\label{sec:exp2}
\subsubsection{Experimental Setup}
To demonstrate that R-LIME learns a highly accurate linear approximation model
in the optimized approximation region,
we conducted a comparison of the local accuracy of explanations
between LIME and R-LIME.
Using the same settings as in \cref{sec:exp_setting},
we randomly sampled 100 instances from the test data of the recidivism dataset
and generated explanations (with $\tau=0.70,0.80,0.90$) using LIME and R-LIME.
We then sampled 10,000 instances within the rectangular region
obtained by R-LIME and calculated the local accuracy of both methods.
\subsubsection{Experimental Results}
\begin{figure}[tbp]
  \centering
  \includegraphics[width=0.55\textwidth]{python/exp2/box_plot}
  \caption[Comparison of Local Accuracy between R-LIME and LIME]{%
    Comparison of local accuracy between LIME and R-LIME.
    R-LIME achieved higher and less variable accuracy compared to LIME
    for all values of $\tau$.
  }\label{fig:box_plot}
\end{figure}
The results are presented in \cref{fig:box_plot},
showcasing the distribution of the accuracy of the linear approximation models
learned by LIME and R-LIME.
R-LIME exhibits higher accuracy compared to LIME for all values of $\tau$.
This suggests that the linear classifiers learned by LIME and R-LIME
differ significantly,
and R-LIME learns a high-accuracy linear classifier adapted to the rectangular region.
Additionally, as $\tau$ increases,
the variability in the accuracy of LIME widens.
This indicates that the linear classifiers learned by LIME may not function
effectively as approximation models depending on how the region is selected.

@@

@@

\subsection{Evaluation through User Studies}
Given the nature of the field of "interpretable machine learning,"
where the goal is to "interpret" machine learning models,
it is challenging to quantitatively compare various methods.
While quantitative experiments were conducted in \cref{sec:exp2},
they mainly demonstrated that the proposed method generates
locally accurate explanations compared to LIME.
The results do not directly prove the high interpretability of the proposed method.

One potential method to quantitatively evaluate "interpretability" is
through user studies.
For example, \cite{ribeiro2018anchors} conducted user studies,
assuming that if users could predict the output of a black-box model
with high accuracy based on provided explanations,
those explanations contained valuable information about the model's behavior.
Similarly,
conducting user studies in this research would be desirable
to evaluate the interpretability of the proposed method.

@@

\subsubsection{Acknowledgements}
% Please place your acknowledgments at
% the end of the paper, preceded by an unnumbered run-in heading (i.e.
% 3rd-level heading).
I would like to express my gratitude to Associate Professor Keigo Kimura
from the Department of Mathematical Science,
Division of Information Recognition,
Graduate School of Information Science and Technology, Hokkaido University.
He provided invaluable guidance,
from setting the research theme to outlining the strategy and content of this study.
Additionally,
I appreciate the diverse insights shared by Professor Mineichi Kudo in the same laboratory.
Finally, I am deeply thankful to all the members of the laboratory
for their helpful advice and collaboration throughout the research process.

>>>>

This paper specifically focuses on \emph{local} \emph{model-agnostic} methods.
For instance, LIME (Local Interpretable Model-agnostic Explanations)
\cite{ribeiro2016why} approximates the complex model's decision boundary
locally using a simpler model around the given focal point.
Although LIME enables users to interpret the local behavior of complex models,
it does not explicitly provide the effective scope of the explanation
\cite{ribeiro2018anchors}.
This limitation can lead users to overgeneralize the knowledge derived from the explanation.
Without a clear indication of the scope of the explanation,
users might mistakenly apply the derived knowledge to instances outside this scope,
potentially leading to a misunderstanding of the black-box model's behavior.

To address this limitation of LIME,
we propose a new method called R-LIME (Ruled LIME),
which interprets complex classifiers in an interpretable region.
R-LIME locally approximates a complex decision boundary linearly
in a rectangular region and maximizes the region as long as the accuracy of the linear
classifier is higher than a specified threshold.
The resulting rectangular region is interpretable for users because it is
expressed as a conjunction of feature predicates.

\section{Related Work}
\begin{figure}[t]
  \centering
  \vspace{0.5cm}
  \begin{subfigure}[t]{0.55\textwidth}
    \centering
    \includegraphics[width=\textwidth]{lime_vs_anchor_exp_a.png}
    \caption{Instances around the focal point.}\label{fig:lime_vs_anchor_exp_a}
    \vspace{0.5cm}
  \end{subfigure}
  \begin{subfigure}[t]{0.45\textwidth}
    \centering
    \includegraphics[width=\textwidth]{lime_vs_anchor_exp_b.png}
    \caption{%
      Explanations by LIME\@.
      They provide contributions of each feature to the output,
      but do not explicitly indicate its scope.
    }\label{fig:lime_vs_anchor_exp_b}
    \vspace{0.5cm}
  \end{subfigure}
  \begin{subfigure}[t]{0.55\textwidth}
    \centering
    \includegraphics[width=\textwidth]{lime_vs_anchor_exp_c.png}
    \caption{%
      Explanations by Anchor.
      They provide the effective scope of the explanation,
      but do not provide details about the influence of each feature.
    }\label{fig:lime_vs_anchor_exp_c}
  \end{subfigure}
  \caption{%
    Explanations by LIME and Anchor
    for an LSTM sentiment prediction model \cite{ribeiro2018anchors}.
  }\label{fig:lime_vs_anchor_exp}
\end{figure}
In this chapter,
we provide an overview of existing research on local model-agnostic explanations,
particularly focusing on studies closely related to our proposed method.

@@

\subsection{Quantitative Evaluation}\label{sec:exp2}
\subsubsection{Experimental Setup}
To demonstrate that R-LIME learns a highly accurate linear approximation model
in the optimized approximation region,
we conducted a comparison of the local accuracy of explanations
between LIME and R-LIME\@.
Using the same settings as in \cref{sec:exp_setting},
we randomly sampled 100 instances from the test data of the recidivism dataset
and generated explanations (with $\tau=0.70,0.80,0.90$) using LIME and R-LIME\@.
We then sampled 10,000 instances within the rectangular region
obtained by R-LIME and calculated the local accuracy of both methods.
\subsubsection{Experimental Results}
\begin{figure}[tbp]
  \centering
  \includegraphics[width=0.55\textwidth]{python/exp2/box_plot}
  \caption[Comparison of Local Accuracy between R-LIME and LIME]{%
    Comparison of local accuracy between LIME and R-LIME\@.
    R-LIME achieved higher and less variable accuracy compared to LIME
    for all values of $\tau$.
  }\label{fig:box_plot}
\end{figure}
The results are presented in \cref{fig:box_plot},
showcasing the distribution of the accuracy of the linear approximation models
learned by LIME and R-LIME\@.
R-LIME exhibits higher accuracy compared to LIME for all values of $\tau$.
This suggests that the linear classifiers learned by LIME and R-LIME
differ significantly,
and R-LIME learns a high-accuracy linear classifier adapted to the rectangular region.
Additionally, as $\tau$ increases,
the variability in the accuracy of LIME widens.
This indicates that the linear classifiers learned by LIME may not function
effectively as approximation models depending on how the region is selected.

@@

\subsection{Computational Complexity}
Most post-hoc explanation methods, including LIME, Anchor, BELLA and R-LIME,
need to query the black-box model multiple times
to get labels for perturbed samples, which can be computationally expensive.
The number of queries required for LIME is $|\mathcal{Z}_p|$,
where $|\mathcal{Z}_p|$ is the number of perturbed samples.

@@

% \subsection{Evaluation through User Studies}
% Given the nature of the field of "interpretable machine learning,"
% where the goal is to "interpret" machine learning models,
% it is challenging to quantitatively compare various methods.
% While quantitative experiments were conducted in \cref{sec:exp2},
% they mainly demonstrated that the proposed method generates
% locally accurate explanations compared to LIME.
% The results do not directly prove the high interpretability of the proposed method.
% 
% One potential method to quantitatively evaluate "interpretability" is
% through user studies.
% For example, \cite{ribeiro2018anchors} conducted user studies,
% assuming that if users could predict the output of a black-box model
% with high accuracy based on provided explanations,
% those explanations contained valuable information about the model's behavior.
% Similarly,
% conducting user studies in this research would be desirable
% to evaluate the interpretability of the proposed method.

@@

% \subsubsection{Acknowledgements}
% Please place your acknowledgments at
% the end of the paper, preceded by an unnumbered run-in heading (i.e.
% 3rd-level heading).
